---
title: "Survival Analysis of Customer Time-to-Churn"
author: "Corey Neff"
date: "6/25/2021"
output:
 github_document:
    html_preview: false
  
---
## Introduction

This is a short project investigating customer time-to-churn of a public dataset. The main goal is to discover the role that subscribing to multiple services at this company plays in probability of churn, but it will also consider customer satisfaction, as well. 

We will start by reading the data, and turning churn value ("Yes", "No") into a factor, while doing some basic visualizations.

```{r}
librarian::shelf(survival, survminer, dplyr, ggplot2)
survdata <- read.csv("survival.csv")


head(survdata, n=5)
survdata$churn_factor <- as.factor(survdata$churn_value)
hist(survdata$months, xlab="Length of Survival Time (Months)", main="Histogram of Survial Time for Customer Churn")
```

## First Kaplan-Meier Curve

Kaplan-Meier Curves are a non-parametric way to vizualize survival data such as these. The Kaplan-Meier survival curve is defined as the probability of "surviving" in a given length of time while considering time in many small intervals. Here, "survival time" is the time-to-churn (i.e. length of time before a cutomer stopped using services, or customer attrition).

```{r pressure, warning=FALSE}
survfit1 <- survfit(Surv(months, churn_value) ~ 1, data = survdata)
basic_plot <- ggsurvplot(survfit1,
           conf.int = T,
           conf.int.fill = "red",
           ylim = c(0.5, 1),
           xlim = c(0, max(survdata$months)),
           xlab = "Time (months)",
           legend = "none",
           title = "Kaplan-Meier Curve for Time to Customer Churn (full sample)",
           ggtheme = theme_bw())
print(basic_plot)
```

## Multiple Services -- Important?

As we can see, change in customer churn appears to be rather linear with respect to time, but this model includes the full sample of data. I'd like to know how the number of services a customer is subscribed to affects survival time. Do customers with less services (i.e. a single service) churn faster than those with more? First we will start with a simple bar chart vizualization to begin to investigate this issue.

```{r}
multiple <- survdata %>%
                  filter(multiple == "Yes") %>%
                  ggplot() + 
                        geom_histogram(aes(x = months, fill  = as.factor(churn_value)), bins = 11) +
                        scale_fill_discrete(name = "Churned?", labels = c("No", "Yes")) +
                        ggtitle("Customers with Multiple Services") +
                        xlab("Time (months)")+
                        theme_bw()

single <-   survdata %>%
                  filter(multiple == "No") %>%
                  ggplot() + 
                        geom_histogram(aes(x = months, fill  = as.factor(churn_value)), bins = 11) +
                        scale_fill_discrete(name = "Churned?", labels = c("No", "Yes")) +
                        ggtitle("Customers with a Single Service")+
                        xlab("Time (months)") +
                        theme_bw()

egg::ggarrange(plots = list(single, multiple), nrow = 1)
```

## Kaplan-Meier Estimator Part Deux 

As we can see, customers subscribing to multiple services tend to survive (i.e. not churn) a much longer time than those subscribed to a single service. We will be able to see this a bit more clearly with another Kaplan-Meier Curve.

```{r}
survfit2 <- survfit(Surv(months, churn_value) ~ multiple, data = survdata)
services_plot <- ggsurvplot(survfit2,
           conf.int = T,
           risk.table = TRUE,
           tables.height = 0.2,
           tables.theme = theme_cleantable(),
           ylim = c(0.5, 1),
           xlim = c(0, max(survdata$months)),
           xlab = "Time (months)",
           legend.title = "Services",
           legend.labs = c("Single", "Multuple"),
           title = "Kaplan-Meier Curve for Time to Customer Churn (full sample)",
           ggtheme = theme_bw())
print(services_plot)
```

## Kaplan-Meier Estimator Part Deux 

So, it seems that customers subscribed to a single service tend to face attrition at a generaly faster rate early-on. That is, early in their patronage, they are much more likely to churn than those subsccribed to multiple services. In the long-run, however, the survival probaility appears to be about the same for both groups.

Now we will begin to investigate a more parametric route -- the Cox Proportional Hazards Model, specifically. Cox models handle multiple covariates much better, which will be useful later. In addition, in a Cox proportional hazards regression model, the measure of effect is the hazard rate, which is the risk of failure (i.e., the risk or probability of suffering the event of interest), given that the participant has survived up to a specific time. Comparing hazard rates allows us to create a hazard ratio, which is similar to an odds ratio.

```{r,warning=FALSE}
coxfit1 <- coxph(Surv(months, churn_value) ~ multiple, data = survdata)
summary(coxfit1)

cox_services <- ggadjustedcurves(coxfit1, variable = "multiple",
                 ylim = c(0.5, 1),
                 ggtheme = theme_bw(),
                 xlab = "Time (months)",
                 legend.title = "Multiple Services",
                 legend.labs = c("Single", "Multuple"),
                 title = 'Survival Curves for Cox Proportional Hazards Model, by Number of "Services"') 
print(cox_services)
```

## Cox Proportional Hazards Model -- Univariate

The Cox distribution yields similar conclusions to the Kaplan-Meier curve before. Generally, people with multiple services have a higher survival rate with respect to time. From the output, we might turn our attention to the exponentiated coefficient of multiple services variable: 0.788. FIn this case, subscribing to multiple services reduces the hazard by a factor of 0.79, or 21%. Therefore, subscribing to multiple services is associated with a good prognosis as far as churn goes. Additionally, we can see that the covariate and model are both significant.

Now that we understands that, generally, subscribing to multiple services corresponds to less churn, I think a solid next step would be to see which services in particular are better in this regard. With that in mind, this next model will include the individual services being subscribed to (security, backup, protection, and support) as well as customer satisfaction, for good measure.

```{r, warning=FALSE}
coxfit2 <- coxph(Surv(months, churn_value) ~ security + backup + protection + support + satisfaction, data = survdata)
summary(coxfit2)
```

## Cox Proportional Hazards Model -- Multivariate

As we can see, the model and all covariates are statistically significant. Additionally, it seems that subscribing to the security service yields the greatest reduction in hazard at a hazard ratio of 0.3482. Even so, it appears that customer satisfaction has the greatest impact on hazard as indicated by it beign the lowest hazard ratio among the included predictors. For that reason, it may be beneficial to see this effect visually. 

```{r, warning=FALSE}
cox_satisfaction <- ggadjustedcurves(coxfit2, variable = "satisfaction",
                 ylim = c(0, 1),
                 ggtheme = theme_bw(),
                 xlab = "Time (months)",
                 title = 'Survival Curves for Cox Proportional Hazards Model, by "Satisfaction"',
                 legend.title = "Product Satisfaction")
print(cox_satisfaction)
```

## Cox Proportional Hazards Model -- Customer Satisfaction

Customer satisfaction seems to have a particularly pronounced (and understandable) effect on customer attrition over time. Customers reporting high satisfaction (i.e. a 5) survive at a rate of what appears to be nearly 100% at the end of the reporting period, while those reporting low satisfaction (i.e. a 1) survive at an abysmal rate -- nearly entirely gone at 70 months. For that reason, the wisest business decision may be to maintain high levels of customer satisfaction; albeit, this is a rather unsurprising takeaway.


